---
title: "Flight_Price_Airline"
author: "Harper Ream"
date: "4/7/2022"
output: github_document
---



```{r setup, include=FALSE}
# install.packages("skimr")
# install.packages("janitor")
# install.packages("lubridate")
# install.packages("zoo")
# install.packages("ggpubr")
```

```{r library}
library(tidyverse)
library(readr)
library(ggplot2)
library(lubridate)
library(skimr)
library(dplyr)
library(zoo)
library(ggpubr)
library(tidyr)
library(stringr)
```

```{r Load theme, include=FALSE}
# theme_set(theme_pubr())
```

```{r Load Dataset}
airline_flights <- read.csv("Clean_Dataset.csv")
```

```{r Load Table}
head(airline_flights)
```

```{r summary}
summary(airline_flights$price)
```


Does price vary with Airlines?


```{r Mean of Airlines}
df_airline_mean <- airline_flights %>% 
  group_by(airline) %>% 
  summarise_at(vars(price), list(airline_mean = mean))
```

```{r mean price per airline}
ggplot(df_airline_mean, aes(x = airline , y = airline_mean, fill = airline)) +
  geom_col() + scale_fill_manual(
    values = c(
      "Air_India" = "#FE9901",
      "AirAsia" = "#E32526",
      "GO_FIRST" = "#094a9a", 
      "Indigo" = "#001B94",
      "SpiceJet" = "#FAD903",
      "Vistara" = "#47143D")) +
    labs(title = "Mean Price per Airline",
         x = NULL,
         y = "Mean Price in Rupee",
         caption = "(based on data from https://www.kaggle.com/shubhambathwal/flight-price-prediction)" ) +
    scale_y_continuous(breaks = seq(1000, 31000, by = 1500),
                       minor_breaks = NULL) +
    theme(plot.title = element_text(face = "bold"))
```

How is the price affected when tickets are bought in just 1 or 2 days before departure?

 
```{r Value range days left}
df_bookings_total <- airline_flights %>%
  filter(airline_flights$days_left %in% 1:50) %>% 
  group_by(days_left) %>% 
  summarise(mean=mean(price))
```

```{r total graph}
total_graph <- ggplot(df_bookings_total) + 
  geom_point(aes(x = days_left, y = mean)) +
  labs(title = "Mean Price vs Days Left Before Departure",
       x = "Days left before Departure",
       y = "Mean Price in Rupee")
total_graph
```

```{r count days left}
hist(airline_flights$days_left)
```

The plane is trying to get more people to fill up the seats in the last day before departure.

```{r df: Price vs days_left and Business}
df_bookings_business <- airline_flights %>%
  filter(airline_flights$days_left %in% 1:50 & 
           #airline_flights$airline == "Vistara" & 
           airline_flights$class == "Business") %>% 
  group_by(days_left) %>% 
  summarise(mean=mean(price))
```

```{r business graph}
business_graph <- ggplot(df_bookings_business) + geom_point(aes(x = days_left, y = mean)) +
  labs(title = "Business Class",
       x = "Days Left Before Departure",
       y = "Mean Price in Rupee")
business_graph
```

```{r df: Price vs days_left and Economy}
df_bookings_economy <- airline_flights %>%
  filter(airline_flights$days_left %in% 1:50 & 
           airline_flights$class == "Economy") %>% 
  group_by(days_left) %>% 
  summarise(mean=mean(price))
```

```{r economy graph}
economy_graph <- ggplot(df_bookings_economy) + geom_point(aes(x = days_left, y = mean)) + 
  labs(title = "Economy Seats",
       x = "Days Left before Departure",
       y = "Mean Price in Rupee")
```


```{r combine business and economy}
figure <- ggarrange(total_graph,
                    ggarrange(business_graph, economy_graph, ncol = 2,
                    labels = c("B", "C")),
                    nrow = 2,
                    labels = "A")
figure
```




Why is the price so low on the last day before departure compared to the 2nd to last day before departure?

GO_FIRST, AirAsia, SpiceJet only have Economy class compared to Business class

